Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-05 ms.tex
fixed nicks tscale definition to something more palatable
fixed abstract

\begin{abstract}
  We calculate steady-state, one-dimensional hydrodynamic profiles of
  hot gas in the nominally quiescent galactic nuclei for a range of
  central black hole masses $M_{\bullet}$, parameterized gas heating
  rates, and observationally-motivated stellar density profiles.  Mass
  is supplied to the circumnuclear medium by stellar winds, while gas
  is heated primarily by stellar winds, supernovae, and black hole
  feedback.  Analytic estimates are derived for the stagnation radius
  (where the radial velocity of the gas passes through zero) and the
  black hole accretion rate, $\dot{M}$, as a function of $M_{\bullet}$ and the gas heating efficiency, the latter of which is
  related to the star-formation history.  We assess the conditions
  under which radiative instabilities develop, separately in the case
  of a single burst of star formation and for the average
  star formation history, as predicted by cosmological simulations, of galaxies hosting black holes of a given mass. The measured nuclear X-ray luminosities, $L_{X}$, of nearby quiescent galactic nuclei and low luminosity AGN are combined with our results for $\dot{M}(M_{\bullet})$ to address whether the nuclei of most nearby galaxies are consistent with accreting in a steady-state, thermally-stable manner.  The maximum accretion rate of a thermally-stable flow is too low to explain the short average growth times of low mass black holes in the local
  universe.  The latter must instead result from gas being fed in from
  large radii, either due to galaxy mergers associated with
  hierarchical growth or thermal instabilities on larger, galactic
  scales.  Our results have implications for attempts to constrain the
  occupation fraction of SMBHs in low mass galaxies using the mean
  $L_{\rm X}-M_{\bullet}$ correlation, as well as applications to other topics, such as the predicted diversity of the
  circumnuclear media encountered by relativistic outflows from
  tidal disruption events.
\end{abstract}

@@

For low-mass black holes, stellar winds are the most important source of heating on
timescales of less than a few tens of millions of years.  Compton
heating comes to dominate at later times due to the higher accretion
rates that accompany the overall decrease in the heating rate, coupled
with the persistently high mass loss rates and gas densities associated with a relatively young
stellar population.  Compton heating dominates at all epochs for
higher mass black holes, with stellar winds and Ia SN competing for
the most significant non-AGN heating source.  As the total heating
rate declines with time, after some point the flow becomes thermally
unstable according to the criterion $(\dot{q}_{\rm heat}/|\dot{q}_{\rm
  rad}|)_{r_{s}} \lesssim 10$ (eq.~[\ref{eq:cooling3}]) {\bf BDM: correct?}, as shown by
dashed lines in Fig.~\ref{fig:vwSourcesImp}.  The thermal stability threshold is
calculated neglecting Compton heating because SMBH feedback may not be
capable of truly stabilizing the flow given its inability to respond
instantaneously to local changes in the properties of the gas.
Thermally unstable flow is seen to be the norm in the single starburst
case, except at very early times, $\tau_{\star} \lesssim 1-3 \times
10^7$ yr.

>>>>

\begin{abstract}
  We calculate steady-state, one-dimensional hydrodynamic profiles of
  hot gas in ``quiescent" galactic nuclei for a range of
  central black hole masses $M_{\bullet}$, parameterized gas heating
  rates, and observationally-motivated stellar density profiles.  Mass
  is supplied to the circumnuclear medium by stellar winds, while gas
  is heated primarily by stellar winds, supernovae, and black hole
  feedback.  Analytic estimates are derived for the stagnation radius
  (where the radial velocity of the gas passes through zero) and the
  black hole accretion rate, $\dot{M}$, as a function of $M_{\bullet}$ and the gas heating efficiency, the latter being related to the star-formation history.  We assess the conditions under which radiative instabilities develop, separately in the case
  of a single burst of star formation and for the average star formation history predicted by cosmological simulations for galaxies hosting black holes of a given mass.  By combining a sample of measured nuclear X-ray luminosities, $L_{X}$, of nearby quiescent galactic nuclei with our results for $\dot{M}(M_{\bullet})$ we show that most nuclei are consistent with accreting in a steady-state, thermally-stable manner.  However, thermally-stable accretion cannot explain the short average growth times of low mass black holes in the local universe, which must instead result from gas being fed in from
  large radii, due either to galaxy mergers or thermal instabilities acting on larger, galactic scales.  Our results have implications for attempts to constrain the occupation fraction of SMBHs in low mass galaxies using the mean
  $L_{\rm X}-M_{\bullet}$ correlation, as well as the predicted diversity of the
  circumnuclear densities encountered by relativistic outflows from
  tidal disruption events.
\end{abstract}

@@

For low-mass black holes, stellar winds are the most important source of heating on
timescales of $\tau_{\star} \lesssim $ few $10^{7}$ years.  Compton
heating comes to dominate at later times due to the higher accretion
rates that accompany the overall decrease in the heating rate, coupled
with the persistently high mass loss rates and gas densities associated with a relatively young
stellar population.  Compton heating dominates at all epochs for
higher mass black holes, with stellar winds and Ia SN competing for
the most significant non-AGN heating source.  As the total heating
rate declines with time, after some point the flow becomes thermally
unstable according to the criterion $(\dot{q}_{\rm heat}/|\dot{q}_{\rm
  rad}|)_{r_{s}} \lesssim 10$ (eq.~[\ref{eq:cooling3}]) {\bf BDM: correct?}, as shown by
dashed lines in Fig.~\ref{fig:vwSourcesImp}.  The thermal stability threshold is
calculated neglecting Compton heating because SMBH feedback may not be
capable of truly stabilizing the flow given its inability to respond
instantaneously to local changes in the properties of the gas.
Thermally unstable flow is seen to be the norm in the single starburst
case, except at very early times, $\tau_{\star} \lesssim 1-3 \times
10^7$ yr.
